---
title: "Week 15"
author: "Sofía Carpio"
date: "4/13/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidytuesdayR)
library(dplyr)
library(ggplot2)
library(ggthemes)
library(lubridate)
library(tidyverse)
library(ggExtra)
library(ggsci)
library(scales)
library(showtext)
library(RColorBrewer)
```

## Week 15 of Tidy Tuesday

This week's data set contained statistics on deaths from indoor air pollution all around the world. A summary of death rates by subregion and continent is shown below. Caveat: I am assuming these are measured in number of deaths for every 100.000 inhabitants, as the data dictionary was not clear.

**Importing the data**

```{r, echo = T, results='hide'}

indoor_pollution <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2022/2022-04-12/indoor_pollution.csv')

View(indoor_pollution)

tuesdata <- tidytuesdayR::tt_load('2022-04-12')

death_ts <- tuesdata$death_timeseries
death_source <- tuesdata$death_source
fuel_gdp <- tuesdata$fuel_gdp
death_full <- tuesdata$death_fuel

View(fuel_gdp)
View(death_full)

country_codes <- readr::read_csv("https://gist.githubusercontent.com/tadast/8827699/raw/f5cac3d42d16b78348610fc4ec301e9234f82821/countries_codes_and_coordinates.csv")

View(country_codes)

```

**Data Wrangling I**

```{r, echo = T, results='hide'}

unique(death_source$Entity)
unique(death_source$Entity[which(is.na(death_source$Code)==TRUE)])

regions <- c("Africa","America","Asia","Europe","Oceania")

Income_level <- c("World Bank High Income","World Bank Low Income","World Bank Lower Middle Income","World Bank Upper Middle Income")

subregion <- c("Caribbean","Central Latin America","North America","Southern Latin America",
               "Central Asia","Central Europe","Central sub-Saharan Africa","East Asia","Eastern Europe",
               "Eastern sub-Saharan Africa","Middle East & North Africa","Southeast Asia","Nordic Region",
               "Southern sub-Saharan Africa", "Western Europe", "Western sub-Saharan Africa",
               "South Asia - World Bank region","Oceania","World")

SDI <- c("High-middle SDI","Low SDI","Low-middle SDI","Middle SDI","High SDI")

SDI_deaths <- death_source %>%
  filter(Entity %in% SDI)

#View(SDI_deaths)

subregion_deaths <- death_source %>%
  filter(Entity %in% subregion)

#View(subregion)

subregion_deaths$continent <- ifelse(subregion_deaths$Entity %in% c("Caribbean","Central Latin America","North America","Southern Latin America"), "Americas",
                                     ifelse(subregion_deaths$Entity %in% c("Central Asia","East Asia","Southeast Asia","South Asia - World Bank region"), "Asia",
                                            ifelse(subregion_deaths$Entity %in% c("Central Europe","Eastern Europe","Western Europe","Nordic Region"),"Europe", 
                                                                                  ifelse(subregion_deaths$Entity %in% c("Central sub-Saharan Africa","Eastern sub-Saharan Africa","Middle East & North Africa","Southern sub-Saharan Africa","Western sub-Saharan Africa"),"Africa",
                                                                                         ifelse(subregion_deaths$Entity == "Oceania","Oceania","World")))))

table(subregion_deaths$Entity, subregion_deaths$continent, useNA = "ifany")

```

**Data Wrangling II**


```{r}

colnames(subregion_deaths)
#"Deaths - Cause: All causes - Risk: Household air pollution from solid fuels - Sex: Both - Age: Age-standardized (Rate)"


colnames(subregion_deaths)[4] <- "Deaths"
subregion_deaths$entity <- factor(subregion_deaths$Entity, levels = c("Caribbean",
                                                                      "Central Latin America","North America",
                                                                      "Southern Latin America",
                                                                      "Central Asia","East Asia","Southeast Asia",
                                                                      "South Asia - World Bank region",
                                                                      "Central Europe","Eastern Europe","Western Europe","Nordic Region",
                                                                      "Central sub-Saharan Africa","Eastern sub-Saharan Africa",
                                                                      "Middle East & North Africa","Southern sub-Saharan Africa",
                                                                      "Western sub-Saharan Africa","Africa","Oceania","World"))

#"Reds","Greens","Blues"
#4,4,4,5,1

reds <- brewer.pal(6,"Reds")
greens <- brewer.pal(6,"Greens")
blues <- brewer.pal(6,"Blues")
purples <- brewer.pal(8,"Purples")
bupu <- brewer.pal(8,"BuPu")
ylorrd <- brewer.pal(5, "YlOrRd")

paleta <- c(reds,greens,blues,purples,"white",bupu, ylorrd)
barplot(rep(1,45),col = paleta)

Americas <- subregion_deaths %>%
  filter(Entity %in% c("Caribbean","Central Latin America","North America","Southern Latin America"))

Asia <- subregion_deaths %>%
  filter(Entity %in% c("Central Asia","East Asia","Southeast Asia","South Asia - World Bank region"))

Europe <- subregion_deaths %>%
  filter(Entity %in% c("Central Europe","Eastern Europe","Western Europe","Nordic Region"))

Africa <- subregion_deaths %>%
  filter(Entity %in% c("Central sub-Saharan Africa","Eastern sub-Saharan Africa",
                       "Middle East & North Africa","Southern sub-Saharan Africa",
                       "Western sub-Saharan Africa","Africa"))


library(showtext)
font_add("hydrophilia", "C:/Users/sofia/Downloads/hydrophilia/Hydrophilia/HydrophiliaIced-Regular.ttf")
font_add("OrelegaOne", "C:/Users/sofia/Downloads/Orelega_One/OrelegaOne-Regular.ttf")

showtext_auto()


```

# The final patchwork plot

```{r}

#titles_sizes 

tit <- 19
sub_cap <- 16
tit_x <- 12
tit_y <- 12

G_Americas <- ggplot(data = Americas, 
            aes(x=Year, y=Deaths,group = entity)) +
  geom_line(aes(color = entity), size = 1) +
  scale_color_manual(values = reds[c(2,4:6)], labels = c("Caribbean","Central Latin\nAmerica","North America","Southern Latin\nAmerica")) +
  scale_x_continuous(breaks = seq(1990,2020,5), expand = c(0,0),
                     labels = seq(1990,2020,5)) +
  theme_bw() +
  labs(title = "Americas",
       #subtitle = "Number of deaths per 100.000 inhabitants",
       y= "Number of deaths per\n100.000 inhabitants",
       color = "Subregion") +
  theme(#plot.background = element_rect(fill = "ivory",
         #                               color = "ivory"),
        plot.margin = margin(rep(15, 4)),
        plot.title = element_text(hjust = 0.5, size = tit, face = "bold", family = "OrelegaOne"),
        plot.subtitle = element_text(hjust = 0, size = sub_cap, color = "black", family = "OrelegaOne"),
        plot.caption = element_text(hjust = 0, size = sub_cap, color = "black", family = "OrelegaOne"),
        axis.title.x = element_text(size = tit_x, color = "black", family = "OrelegaOne"),
        axis.title.y = element_text(size = tit_y, color = "black", family = "OrelegaOne"),
        legend.text = element_text(family = "OrelegaOne"),
        legend.title = element_text(family = "OrelegaOne"))
        #axis.text.x = element_text(color = "black", angle = 90))
  
G_Asia <- ggplot(data = Asia, 
            aes(x=Year, y=Deaths,group = entity)) +
  geom_line(aes(color = entity), size = 1) +
  scale_color_manual(values = greens[c(2,4:6)], labels = c("Central Asia","East Asia","Southeast Asia","South Asia -\nWB region")) +
  scale_x_continuous(breaks = seq(1990,2020,5), expand = c(0,0)) +
  theme_bw() +
  labs(title = "Asia",
       #subtitle = "Number of deaths per 100.000 inhabitants",
       y= "Number of deaths per\n100.000 inhabitants",
       color = "Subregion") +
  theme(plot.margin = margin(rep(15, 4)),
        plot.title = element_text(hjust = 0.5, size = tit, face = "bold",family = "OrelegaOne"),
        plot.subtitle = element_text(hjust = 0, size = sub_cap, color = "black",family = "OrelegaOne"),
        plot.caption = element_text(hjust = 0, size = sub_cap, color = "black",family = "OrelegaOne"),
        axis.title.x = element_text(size = tit_x, color = "black",family = "OrelegaOne"),
        axis.title.y = element_text(size = tit_y, color = "black",family = "OrelegaOne"),
        axis.text = element_text(color = "black"),
        legend.text = element_text(family = "OrelegaOne"),
        legend.title = element_text(family = "OrelegaOne"))

G_Europe <- ggplot(data = Europe, 
            aes(x=Year, y=Deaths,group = entity)) +
  geom_line(aes(color = entity), size = 1) +
  scale_color_manual(values = blues[c(2,4:6)]) +
  scale_x_continuous(breaks = seq(1990,2020,5), expand = c(0,0),
                     labels = seq(1990,2020,5)) +
  theme_bw() +
  labs(title = "Europe",
       #subtitle = "Number of deaths per 100.000 inhabitants",
       y= "Number of deaths per\n100.000 inhabitants",
       color = "Subregion") +
  theme(plot.margin = margin(rep(15, 4)),
        plot.title = element_text(hjust = 0.5, size = tit, face = "bold",family = "OrelegaOne"),
        plot.subtitle = element_text(hjust = 0, size = sub_cap, color = "black",family = "OrelegaOne"),
        plot.caption = element_text(hjust = 0, size = sub_cap, color = "black",family = "OrelegaOne"),
        axis.title.x = element_text(size = tit_x, color = "black",family = "OrelegaOne"),
        axis.title.y = element_text(size = tit_y, color = "black",family = "OrelegaOne"),
        legend.text = element_text(family = "OrelegaOne"),
        legend.title = element_text(family = "OrelegaOne"))
        #axis.text.x = element_text(color = "black", angle = 90))

G_Africa <- ggplot(data = Africa, 
            aes(x=Year, y=Deaths,group = entity)) +
  geom_line(aes(color = entity), size = 1) +
  scale_color_manual(values = bupu[c(2,4,5,7,8)], labels= c("Central\nsub-Saharan A.","Eastern\nsub-Saharan A.",
                       "Middle East &\nNorth Africa","Southern\nsub-Saharan A.",
                       "Western\nsub-Saharan A.")) +
  scale_x_continuous(breaks = seq(1990,2020,5), expand = c(0,0),
                     labels = seq(1990,2020,5)) +
  theme_bw() +
  labs(title = "Africa",
       #subtitle = "Number of deaths per 100.000 inhabitants",
       y= "Number of deaths per\n100.000 inhabitants",
       color = "Subregion") +
  theme(plot.title = element_text(hjust = 0.5, size = tit, face = "bold",family = "OrelegaOne"),
        plot.subtitle = element_text(hjust = 0, size = sub_cap, color = "black",family = "OrelegaOne"),
        plot.caption = element_text(hjust = 0, size = sub_cap, color = "black",family = "OrelegaOne"),
        axis.title.x = element_text(size = tit_x, color = "black",family = "OrelegaOne"),
        axis.title.y = element_text(size = tit_y, color = "black",family = "OrelegaOne"),
        legend.text = element_text(family = "OrelegaOne"),
        legend.title = element_text(family = "OrelegaOne"))
        #axis.text.x = element_text(color = "black", angle = 90))
  

G_Oceania <- ggplot(data = subregion_deaths[subregion_deaths$Entity == "Oceania",], 
            aes(x=Year, y=Deaths,group = entity)) +
  geom_line(aes(color = entity), size = 1) +
  scale_color_manual(values = ylorrd[2]) +
  scale_x_continuous(breaks = seq(1990,2020,5), expand = c(0,0),
                     labels = seq(1990,2020,5)) +
  scale_y_continuous(limits = c(0,250)) +
  theme_bw() +
  labs(title = "Oceania",
       #subtitle = "Number of deaths per 100.000 inhabitants",
       y= "Number of deaths per\n100.000 inhabitants",
       color = "Subregion") +
  theme(plot.title = element_text(hjust = 0.5, size = tit, face = "bold",family = "OrelegaOne"),
        plot.subtitle = element_text(hjust = 0, size = sub_cap, color = "black", family = "OrelegaOne"),
        plot.caption = element_text(hjust = 0, size = sub_cap, color = "black", family = "OrelegaOne"),
        axis.title.x = element_text(size = tit_x, color = "black", family = "OrelegaOne"),
        axis.title.y = element_text(size = tit_y, color = "black",family = "OrelegaOne"),
        legend.position = "none")
        #axis.text.x = element_text(color = "black", angle = 90))

G_World <- ggplot(data = subregion_deaths[subregion_deaths$Entity == "World",], 
            aes(x=Year, y=Deaths,group = entity)) +
  geom_line(aes(color = entity), size = 1) +
  scale_color_manual(values = "grey") +
  scale_x_continuous(breaks = seq(1990,2020,5), expand = c(0,0),
                     labels = seq(1990,2020,5)) +
  theme_bw() +
  labs(title = "World",
       #subtitle = "Number of deaths per 100.000 inhabitants",
       y= "Number of deaths per\n100.000 inhabitants",
       color = "Subregion") +
  theme(plot.title = element_text(hjust = 0.5, size = tit, face = "bold",family = "OrelegaOne"),
        plot.subtitle = element_text(hjust = 0, size = sub_cap, color = "black",family = "OrelegaOne"),
        plot.caption = element_text(hjust = 0, size = sub_cap, color = "black",family = "OrelegaOne"),
        axis.title.x = element_text(size = tit_x, color = "black",family = "OrelegaOne"),
        axis.title.y = element_text(size = tit_y, color = "black",family = "OrelegaOne"),
        legend.position = "none")
        #axis.text.x = element_text(color = "black", angle = 90))
  
library(patchwork)

G_total <- (G_Africa + G_Asia + G_Americas)/(G_Europe + G_Oceania + G_World)

G_final <- G_total + plot_annotation(
  title = "Deaths from indoor air pollution by continent between 1990 and 2019",
  subtitle = "The rate of deaths from indoor pollution have experienced a decline in all regions of the world since 1990, with the exception of a small increase during the early 1990s in\nEastern Europe, Southern sub-Saharan Africa and Central Asia. Number of deaths per 100.000 inhabitants remain above 100 in Central, Eastern and Western sub-Saharan Africa and Oceania.",
  caption = "Source: Our World in Data | Plot: Sofía Carpio @Sofl77",
  theme = theme(plot.title = element_text(size = 30))) &
  theme(text = element_text('OrelegaOne'))

G_final

```





